---
title: "Cycle 2"
author: "Damian Wollny"
output:
  html_document:
    theme: paper
    df_print: kable
    code_folding: hide
date: '`r format(Sys.Date(), "%B %d, %Y")`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = FALSE, message = FALSE)
options(width = 12)
```

```{r}
library(tidyverse)
theme_set(theme_classic(base_size=22))
```

## Course content
 
### Get the data

This time we will get data that is a bit more of a pain to work with and needs some cleaning up before we get started

<u>Things to learn:</u>

 - 'read.csv()' -> 'sep' argument, 'skip' argument
 - string manipulation with stringr -> e.g. get rid of the first/last character, split strings by character...
 

### Manipulation of table

Here the goal would be to learn more ways of data manipulation

<u>Things to learn:</u>

 - for loop and if statements -> not sure how to implement but definately something they need to know
 - combine tables -> bind_cols(), bind_rows(), left_join() etc
 - summarize data -> group_by() and summarize()

### Plots

Now the plots should get more suffisticated and really demonstrate how powerful ggplot2() is

<u>Things to learn:</u>

 - different geoms (this can be quite extensive)
 - facets (if they haven't already)
 - logscale the axes
 - tweaking the plots -> labelling datapoints, choosing color paletts...

## Homework

### Homework types / Goal of homework

One idea for the 2nd cycle would be that we demonstrate the principles using e.g. iris and let them do the real thing with the interesting data at home

