---
title: "Cycle 1"
author: "Damian Wollny"
output:
  html_document:
    theme: paper
    df_print: kable
    code_folding: hide
date: '`r format(Sys.Date(), "%B %d, %Y")`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = FALSE, message = FALSE)
options(width = 12)
```

```{r}
library(tidyverse)
theme_set(theme_classic(base_size=22))
```

## Course content

### Introduce the RStudio environment

Here, the idea is to learn what all the windows and buttons mean

<u>Things to learn:</u>
 
 - Gobal environment, console, terminal, "code window"
 - save a simple script file?
 - markdown language: "new file" -> "markdown" -> "Knit to HTML" -> look at what code ouput looks like
 
### Get the data

First we will try to get the data from the source and try to load it into R

<u>Things to learn:</u>

 - clicking for millenials -> explain what functions, libraries etc. are
 - what does 'read.csv()' mean? -> what is a function?
 - introduce ?help
 - use tab for autocompletion
 - assign variable for the new table (= vs <-)
 - when do you use quotations when assigning variables -> Difference between x <- 2 and x <- "2" -> explain variable types (character, numeric, factor etc)
 - parantheses and quotations must be closed -> common error
 - look at the data with View(), head(), tail(), str() functions
 - comment code with "#"
 - concept of working directory / paths for files / changing directory

### Simple manipulation of table

<u>Things to learn:</u>

 - we need tidyverse -> what is a library? how to install? how to load?
 - concept of pipe
 - clean up table: dplyr basics - introduce filter(), select(), rename() - introduce logical operators
 - read.csv() function (?)

### Simple plots

<u>Things to learn:</u>

 - grammar of graphics (ggplot)
 - explain principle of adding layers
 - start with geom_point (easiest)
 - explain aesthetic mappings -> adding more dimensions to 2-dimensional plot
 - play with color, size, shape, alpha... -> maybe a good point to introduce stack overflow? let them google how to change the point color in a ggplot
 - explain facets (??)

## Homework

### Homework types / Goal of homework

 - "Übung macht den Meister" -> let them repeat the things the things they learned with slight tweeks
 - Why does this code not work? -> Fix broken code
