---
title: "R course - WS22/23"
author: "Damian Wollny"
output:
  html_document:
    theme: paper
    df_print: kable
date: '`r format(Sys.Date(), "%B %d, %Y")`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
theme_set(theme_classic(base_size=18))
```

# A cross-country database of COVID-19 testing

**Paper**: <https://www.nature.com/articles/s41597-020-00688-8#Sec7>

**Data**: <https://github.com/owid/covid-19-data/tree/master/public/data>

<u>About the paper</u>:

"data can only be meaningfully interpreted alongside an accurate understanding of the extent of virus testing in different countries"

"The database is updated regularly"

"We rely on figures published in official sources, including press releases, government websites, dedicated dashboards, and social media accounts of national authorities"



## Data exploration

```{r}
# load data
cov2_data <- read.csv(file = "~/PI/Lectures/R_course/1st_cycle/owid-covid-data.csv") %>% as_tibble()
```

### Does the booster help?

Teaching problem: Vaccination is a sensitive topic

```{r, fig.width=12, fig.height=10}
cov2_data %>% 
  filter(continent == "Europe") %>% 
  mutate(date = as.Date(date)) %>% 
  select(location, date, icu_patients_per_million, total_boosters_per_hundred) %>% 
  drop_na() %>%
  ggplot(aes(x = date, y = icu_patients_per_million, color = total_boosters_per_hundred)) +
  geom_point() +
  scale_x_date(date_breaks = "3 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  facet_wrap(~location, scales = "free_y", ncol = 3) +
  labs(title = "Booster effect")
```

# Are things getting better?

Teaching problem: 'date' column needs to be converted using "as.Date". Comes a bit too early

```{r, fig.width=10}
test <- c("Netherlands", "Germany", "Denmark", "United Kingdom" , "Slovenia")

cov2_data %>% 
  filter(continent == "Europe") %>% 
  filter(location %in% test) %>% 
  mutate(date = as.Date(date)) %>% 
  select(location, date, new_cases_per_million, new_deaths_per_million ) %>% 
  drop_na() %>%
  ggplot(aes(x = new_deaths_per_million, y = new_cases_per_million, color = date)) +
  geom_point() +
  #scale_x_date(date_breaks = "3 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  facet_wrap(~location, scales = "free_y", ncol = 3) +
  labs(title = "More infected, less dead")

```

# Does the number of cases depend on the population density?

Teaching problem: Plot might be too complicated in the sense that there are many concept you need to learn for it to be informative

```{r}
cov2_data %>% 
  filter(date == "2022-09-27") %>% 
  filter(continent == "Europe") %>% 
  select(location, date, population_density, total_cases) %>% 
  ggplot(aes(x = population_density, y = total_cases, label = location)) +
  geom_point() +
  geom_label() +
  scale_y_continuous(trans = "log10") +
  scale_x_continuous(trans = "log10")
```

# Does the number of cases depend on the population age?

This could be interesting

```{r}
cov2_data %>% 
  filter(date == "2022-09-27") %>% 
  select(location, aged_65_older, total_deaths_per_million, continent) %>% 
  ggplot(aes(x = aged_65_older, y = total_deaths_per_million, color = continent)) +
  geom_point(size = 3) +
  scale_y_continuous(trans = "log10")
```

# Are the rich countries dying less?

Teaching problem: Difficult to interpret. Could very well be a 'missing data'/'insufficient testing' problem

```{r}
cov2_data %>% 
  filter(date == "2022-09-27") %>% 
  #filter(continent == "Europe") %>% 
  select(location, date, total_deaths_per_million, gdp_per_capita) %>% 
  ggplot(aes(x = gdp_per_capita, y = total_deaths_per_million, label = location)) +
  geom_point() +
  #geom_label() +
  scale_y_continuous(trans = "log10") +
  scale_x_continuous(trans = "log10")
```

